---
title: 'Practical Machine Learning: Course Project'
author: "Michelle Jaeger"
date: "April 25, 2015"
output: html_document
---

## Introduction
This project uses the Weight Lifting Exercise Dataset, generously made available here: http://groupware.les.inf.puc-rio.br/har). To obtain this data, six participants performed bicep curls in five different fashions. Class A corresponds to the specified execution of the exercise, while the four other classes correspond to common mistakes. Various accelerometer reading variables are used to predict the classe variable.

### Read in Data
```{r}
training <- download.file("https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv", destfile="training.csv", method="curl")
testing <- download.file("https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv", destfile="testing.csv", method="curl")
training_data <- read.csv("training.csv")
testing_data <- read.csv("testing.csv")
```

### Libraries
```{r}
library(caret)
library(randomForest)
```

### Data Cleaning
``` {r}
# Remove variables with near zero variance
nzv <- nearZeroVar(training_data)
training_data <- training_data[,-nzv]
# Remove variables with NAs
filtered_training_data <- training_data[, colSums(is.na(training_data)) == 0]
# Remove irrelevant variables
filtered_training_data <- filtered_training_data[ , 8:59]
```
Columns with near zero variance and/or missing data were removed, and columns for variables that did not seem relevant (names, dates, times, etc.) were removed.

### Cross Validation
``` {r}
set.seed(1)
inTrain <- createDataPartition(y = filtered_training_data$classe, p = .8, list=F)
train <- filtered_training_data[inTrain,]
test <- filtered_training_data[-inTrain,]
cross_validated <- trainControl(method="cv", number=4)
```
A seed was set for reproducibility, the training data was divided into training data and testing data (80% training and 20% testing), and a trainControl variable was created to be used in creating the model.

### Build the model using Random Forest
``` {r}
modFit <- randomForest(classe ~., data=train, trControl=cross_validated)
prediction <- predict(modFit, test)
result <- confusionMatrix(prediction, test$classe)
result
```
The confusion matrix above details the performance of the algorithm. The Accuracy statistic is .995, indicating an **out of sample error** of around .5%.

## Predicting Results on the original Test set
``` {r}
predictionResult <- predict(modFit, testing_data)
predictionResult
```
